### 2025-08-14 paper/paper.tex
latex stuff

\documentclass[12pt]{amsbook}
\usepackage[utf8]{inputenc}
\usepackage[
backend=biber,
style=numeric,
maxbibnames=99,
sorting=ynt
]{biblatex}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{tikz}
\usepackage{standalone}
\usepackage{svg}
\usepackage{gensymb}
\usepackage{float}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{graphicx, animate}
\usepackage{caption, subcaption}
\usepackage{pgffor}
\usepackage{mathtools}
\usepackage{array}
\usepackage{cleveref}
\addbibresource{bibliography.bib}
\usepackage{hyperref,thmtools}
%\usepackage{chngcounter}

\usetikzlibrary{decorations.markings}
\usetikzlibrary{arrows,automata}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta,positioning}

\counterwithin{figure}{chapter}
\counterwithin{section}{chapter}
\counterwithin{table}{chapter}
\counterwithin{subsection}{section}

\linespread{1.5}
\let\cleardoublepage\clearpage
% Override ugly default link
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = blue, %Colour for external hyperlinks
  linkcolor    = blue, %Colour of internal links
  citecolor   = blue    %Colour of citations
}

\begin{document}
\begin{chapter}{Methodology}

        % Goal is to predict next time steps
        We are given a sequence of graphs, and our goal is to predict the next graph in the sequence. To achieve this, we reinterpret our temporal network as a sequence of points in a latent space; that is, a space in which similar vertices (ones with many of the same neighbours) are mapped to a similar point in space.
        We then train a Neural Ordinary Differential Equation (NODE) to approximate the rate of change of the points in this latent space. With this numerical equation, we can then use symbolic regression techniques to find a functional equation that matches the numerical equation. Using this process we discover an interpretable function that governs the evolution of a temporal network.

    \begin{section}{Singular Value Decomposition and Random Dot Product Graphs}
        SVD Calculated with XXX Arpack XXX Julia
        Each system has 2 singular values
        Aligning The L and R matrices is important (The SVDs can each be fliped)
        We use an initial vector to ensure that the first eigenvector is oriented the same for each time step
        Then calculate a transformation for the subsequent eigenvectors, which is used for training loss and prediction.
        L and R are related by [1 0; 0 -1] as the network is not directed (IE the matrix is symetrical) 
        We only model L and transform it for analysis
        Reconstruct the RDPG L*(L*transform)'
    \end{section}

    \begin{section}{Neural Ordinary Differential Equation}
        Used Julia SciML ecosystem
        Basic copy pasted code 
        Modelling the whole state
        Loss function is abs2 sum Data-pred + disabove 1 and below 0 (L*(L*transform)' )
        NN structure 3 layers 256, 128, 128 celu activation function
    \end{section}

    % Symbolic Regression Explaination

    \begin{section}{Symbolic Regression}
        ???
    \end{section}

    \begin{section}{Reconstructing the Temporal Network}
        XXX Is this necessary XXX

        $u$ is the first row of the matrix $\hat L$. By taking $u\hat R'$, this is equivalent to finding $\hat A = \hat L \hat R'$ and looking at $\hat A_{1\cdot}$; to recover a final prediction of the edges of $u$, we take:
        \begin{align}
            q &= u \hat R'
        \end{align}
        This gives a vector $q$ of probabilities of connections between the target node and the rest of the network. As the results can be interpreted as a probability, the most likely graph is given by setting any entry with a value of $0.5$ or higher to one (predict that the edge exists), and setting any entry with a value of less than $0.5$ to zero (predict that an edge does not exist). However, this may not provide the best fit in terms of the accuracy of edges, and techniques to find the best threshold exist (for more information see \Citeauthor{strydomEtAl}\cite{strydomEtAl})
    \end{section}
\end{chapter}

\end{document}